Honeybees vary collective decision making in different landscapes. 

### 2022-07-27 analysis/manuscript/paper.Rmd
updaed figs for updated version

```{r, fig.cap="Figure 2. **Simulating honey bee foraging**. In our simulation model with scouting only (a), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality). When foraging with recruitment (b) foragers continue to identify resources in scouting trips (white lines) and convey this information on the dance floor (brown disc) where foragers can sample dances reporting on scouting and recruiting trips and follow these directions (yellow lines). (c) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips. Note the difference in the shape of the distributions. The scout distribution is best fit by an exponential (grey fit line), the recruit distribution by a Rayleigh distribution (black line)."}
knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
```

@@

```{r, fig.cap="Figure 3. **The rationale of the foraging model**. The distances of resources encountered by scouts are distributed exponentially (a). These dances are advertised on the dance floor (b). Dances for resources that are closer or higher in quality are repeated more often (c). As a consequence, dances for more profitable resource are over-represented and sampling foragers are biased to the more profitable resources (d). After successfully visiting advertised resources, recruits also dance for them leading to further amplification of this bias towards the most profitable resource in the vicinity of the hive (e).  The distances of recruiting trips are than distributed through a Rayleigh distribution (f). The distances reported on the dance floor are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
```

@@

For 16 out of 20 study hives, the collective model provided a better description of the data than the individual model (Fig. 4a). In the other 4 sites, despite the collective model having the higher maximum likelihood, the individual model had a higher AIC value and so is more parsimonious (Supplementary Table 1). In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see Materials and Methods) to the empirical waggle run durations (Fig. 4b), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig. 4b). The sites shown in Figs 4c-d are representative examples showing the model fits where the individual (Fig. 4c) and collective (Fig. 4d) models fit best. Note the closeness of the fit to the data, illustrating the overall quality of the model description.

```{r, fig.cap="Figure 4. **The honey bee foraging model fitted to data from 20 hives**.  (a) Location of study hives in Southern England, shaded area in the main plot indicates Greater London. For 16 hives for the collective foraging model provided best explanation (black circles) or for 4 hives the individual search model provided the best explanation(red triangles) as indicated by lowest AIC score. (b). Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site. The p-value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data, the red dashed line marks the significance threshold of 0.05. For values exceeding the threshold there is no statistically significant difference between the model and the data, indicating the model provides a good fit. For the hive in (c) the individual model (blue line) provided a better fit than the collective foraging model (red line). For the hive in (d) the collective foraging model (red line) provided a better fit than the individual model (blue line). Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
knitr::include_graphics("../results/figures/manuscript_figures/fig4.png")
```

@@

Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips, $p$, for each site. Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use. We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see Materials and Methods). We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments. As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis. Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material).

In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in the proportion of scouts (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig. 5a). This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig. 5b). These land-use types maintain a significant correlation with the first principal component over the jackknifed PLS (Fig. 5b), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust. As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases. Arable land in the UK is typically considered nutritionally poor for bees (note that oilseed rape fields were not included within this category), while there is evidence to suggest that the residential areas that were captured within the “built-up” category are forage-rich hotspots, typically supporting relatively high bee diversity and abundance within gardens [@Baldock2019].

@@

```{r, fig.cap="Figure 5. **Collective foraging correlates with land-use**.  The proportion of scouts for each site against the first principal component derived from a Partial Least Squares analysis of land-use type. Beta regression shows the relationship (black line) between first principal component and the proportion of scouts, with 95% CI shown by the grey shaded area for agri-rural (a) and urban landscapes (c). The correlations between first principal component and each land-use type are shown for agri-rural (b) and urban landscapes (d). Correlations outside the shaded area significantly contribute to the first principal component. Colours correspond to the land use as shown in maps (circular insets) for selected sites. NAUMGS (resp. NAMGS) stands for non-agricultural unmanaged (resp. managed) green space."}
knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
```

In the urban environment, our findings were more equivocal. The first principal component explained ~73% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig. 5c). This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected if the proportion of scouts decreases when rich forage is available, since such land includes a high proportion of gardens. However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig. 2.), and is not a significant contributor to the variance in the loadings (Fig. 5c.). Furthermore, the first principal component also correlates positively with parks, allotments and cemeteries (8% of land cover; also railways but these constitute ~1% of land cover; Table 1), which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig. 3.).

# Discussion New

@@

As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010]. As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation. After analysis the data was back transformed to the original values for the plots in Fig 5.

>>>>

```{r, fig.cap="Figure 2. **Simulating honey bee foraging**. In our simulation model with scouting only (A), foragers leave the hive on a search path (white lines) and continue until they encounter a resource (circles, colours indicate different resource quality). When foraging with recruitment (B) foragers continue to identify resources in scouting trips (white lines) and convey this information on the dance floor (brown disc) where foragers can sample dances reporting on scouting and recruiting trips and follow these directions (yellow lines). (C) Complementary cumulative frequencies of foraging distances reported from scouting and recruit trips. Note the difference in the shape of the distributions. The scout distribution is best fit by an exponential (gray fit line), the recruit distribution by a Rayleigh distribution (black line)"}
knitr::include_graphics("../results/figures/manuscript_figures/fig2.png")
```

@@

```{r, fig.cap="Figure 3. **The rationale of the foraging model**. The distances of resources encountered by scouts are distributed exponentially (A). These dances are advertised on the dance floor (B). Dances for resources that are closer or higher in quality are repeated more often (C). As a consequence, dances for more profitable resource are over-represented and sampling foragers are biased to the more profitable resources (D). After successfully visiting advertised resources, recruits also dance for them leading to further amplification of this bias towards the most profitable resource in the vicinity of the hive (E). The distances of recruiting trips are than distributed through a Rayleigh distribution (F). Recruits also report the locations on the dance floor (B) and repeat their runs more often depending on the profitability of the location (C), leading to a distribution of durations of recruit dances (G). By taking together the dance distributions for scouts (D) and recruits (G) the distributions of all dances on the dance-floor can be found. The distances reported on the dance floor this are a mixture of the scout and recruiting trips and can be calculated from the distance distributions of the scouting and recruiting trips, taking the reporting bias into account (see Materials and Methods for detail)."}
knitr::include_graphics("../results/figures/manuscript_figures/fig3.png")
```

@@

For 16 out of 20 study hives, the collective model provided a better description of the data than the individual model (Fig. 4A). In the other 4 sites, despite the collective model having the higher maximum likelihood, the individual model had a higher AIC value and so is more parsimonious (Supplementary Table 1). In all but one site, the collective model had a good fit (using a Kolmogorov-Smirnov statistic of $p > 0.05$, see Materials and Methods) to the empirical waggle run durations (Fig. 4B), whereas the individual model was significantly different to the observed data in 8 sites (Kolmogorov-Smirnov statistic $p < 0.05$, Fig. 4b). The sites shown in Figs 4C-D are representative examples showing the model fits where the individual (Fig. 4C) and collective (Fig. 4D) models fit best. Note the closeness of the fit to the data, illustrating the overall quality of the model description.

```{r, fig.cap="Figure 4. **The honey bee foraging model fitted to data from 20 hives**. (A) Location of study hives in Southern England, shaded area in the main plot indicates Greater London. For 16 hives for the collective foraging model provided best explanation (black circles) or for 4 hives the individual search model provided the best explanation (red triangles) as indicated by lowest AIC score. (B). Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site. The p-value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data, the red dashed line marks the significance threshold of 0.05. For values exceeding the threshold there is no statistically significant difference between the model and the data, indicating the model provides a good fit. For the hive in (C) the individual model (blue line) provided a better fit than the collective foraging model (red line). For the hive in (D) the collective foraging model (red line) provided a better fit than the individual model (blue line). The typical “hump” in the distribution in (D) which is indicative of contribution of recruitment dances (compare to Fig 3H). Panels show the compliment cumulative frequencies with binned frequency distributions as inset."}
knitr::include_graphics("../results/figures/manuscript_figures/fig4.png")
```

@@

Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of recruit trips, $1-p$, for each site. Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use. We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see Materials and Methods). We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see Materials and Methods) to determine the principal components that represent combinations of land-use types which explained the most variation in waggle dance use within agri-rural and urban environments. As for one of our sites neither model provided a plausible description, this site was removed from the PLS analysis. Due to our small sample size (10 urban and 9 agri-rural sites), we used jackknife resampling to evaluate the robustness of our results to influential points (see Materials and Methods, Supplementary Material).

In the agri-rural environments the first principal component is a combination of land use types that explained ~73% of the variation in waggle dance use (beta regression: $R^2 = 0.73$, $\phi$ = 4.9, p < 0.05, Fig. 5A). This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage); note that it also correlates negatively with non-agricultural unmanaged green space and water, but together these represent less than 3% of land-use (Table 1; Fig. 5B). These land-use types maintain a significant correlation with the first principal component over the jackknifed PLS (Fig. 5B), with the exception of non-agricultural unmanaged green space which sits on the boarder, indicating the results are robust. As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases. Arable land in the UK is typically considered nutritionally poor for bees (note that oilseed rape fields were not included within this category), while there is evidence to suggest that the residential areas that were captured within the “built-up” category are forage-rich hotspots, typically supporting relatively high bee diversity and abundance within gardens [@Baldock2019].

@@

```{r, fig.cap="Figure 5. **Collective foraging correlates with land-use**. Estimated waggle dance use for each site against the first principal component derived from a Partial Least Squares analysis of land-use type. Beta regression shows the relationship (black line) between first principal component and waggle dance use, with 95% CI shown by the grey shaded area for agri-rural (A) and urban landscapes (C). The correlations between first principal component and each land-use type are shown for agri-rural (B) and urban landscapes (D). Correlations outside the shaded area significantly contribute to the first principal component. Colours correspond to the land use as shown in maps (circular insets) for selected sites. NAUMGS (resp. NAMGS) stands for non-agricultural unmanaged (resp. managed) green space"}
knitr::include_graphics("../results/figures/manuscript_figures/fig5.png")
```

In the urban environment, our findings were more equivocal. The first principal component explained ~73% of the variance in waggle dance use (beta regression: $R^2$ = 0.73, $\phi$ = 10.4, p < 0.05, Fig. 5C). This correlated negatively with the dominant land-use type by coverage, sparse residential (land coverage ~35%; Table 1), as would be expected since such land includes a high proportion of gardens. However, this relationship is very sensitive to site removal through the jackknife sampling (Supplementary Fig. 2.), and is not a significant contributor to the variance in the loadings (Fig. 5D.). The first principal component also correlates positively with railways but these constitute ~1% of land cover (Table 1). Furthermore, the first principal component correlates negatively with parks, allotments and cemeteries which would typically be considered forage-rich for honeybees [@Baldock2019] (although note that the amount of land cover for this component is limited and the jackknifed PLS identified substantial variation in the loadings identified for the first principal components; Supplementary Fig. 3.).

# Discussion

@@

As our estimates of waggle dance use is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010]. As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}$, as outlined in the betareg package documentation. After analysis the data was back transformed to the original values for the plots in Fig 5.
